---
title: "Richmond Network Complete Calibration Framework"
author: "Audrey Smith"
date: "7/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(httr)
library(jsonlite)
library(dplyr)
library(lubridate)
library(ggplot2)
library(mailR)
library(stringr)
library(CVXR)
```

### The script is designed to:
1. Read in daily, 60-minute air quality monitoring data from Richmond Network, previously downloaded with Python script

2. Call regulatory API and download data for applicable proxy station
   API documentation: https://docs.airnowapi.org/

3. Compare data to proxy data for drift detection
3.a. Run Kolmogorov Smirnov test to compare three-day rolling average of each monitor to that of the BAAQMD data
    3.a.i. A flag should be created for each time KS statistic goes below 0.05
        3.a.ii. Create separate column for running count of flag

3.b. Run deviance test to see if mean and variance of our data differ substantially with those of the proxy data
    3.b.i. A separate flag should be created for each of these, with a running flag column 

4. After five consecutive days of flags for any parameter, recalibrate data
    4.a. Calibration uses mean and variance of our data, proxy data to create new slope and offset for O3
    4.b. Calibration uses mean and variance approach, as well as minimizing Kullback Leibler Divergence between data and proxy distributions
    4.c. Data should be re-calibrated back to first day of flag, using eight days of data for slope and offset calculation
        
5. Output - running log of calibration parameters for each monitor for NO2 and O3

### Proxy Data Collection
```{r}
get_proxy <- function(start_proxy, end_proxy, pollutant){
  request_url = paste0('http://www.airnowapi.org/aq/data/?startDate=', start_proxy, 'T00&endDate=', end_proxy, 'T23&parameters=OZONE,PM25,NO2&BBOX=-123.201658,37.177199,-121.366941,38.530363&dataType=C&format=application/json&verbose=1&nowcastonly=0&includerawconcentrations=1&API_KEY=C05358E3-5508-4216-A03E-E229E0368B7E')
  request_call <- GET(url = request_url)
  request_json <- content(request_call, as = 'text', type = NULL, encoding = 'UTF-8')
  request_df <- arrange(fromJSON(request_json, simplifyDataFrame = TRUE), desc(UTC))
    
  # Clean response data
  reg_data <- mutate(request_df, timestamp = paste0(substr(UTC, 1, 10), ' ', substr(UTC, 12, 16), ':00'), RawConcentration = ifelse(RawConcentration == -999, Value, RawConcentration))
  
  # Generate random noise to approximate continuous distribution
  rand_noise <- runif(nrow(reg_data), .00001, .00009)
  reg_data$proxy_rand <- reg_data$RawConcentration + rand_noise
  
  # Clean col names and filter to pollutant of interest
  reg_data <- dplyr::select(reg_data, -c(UTC, Value, AgencyName, FullAQSCode, IntlAQSCode)) %>% 
      rename('lat'='Latitude', 'lon'='Longitude', 'modality'='Parameter', 'units'='Unit', 'proxy_raw'='RawConcentration', 'proxy_site' = 'SiteName') %>%
      filter(modality == pollutant)
  
  return(reg_data)
}
```

### Drift Detection and Recalibration
```{r}
calibrate_monitors <- function(start_72, pollutant){
  
  end_72 <- start_72+2
  
  ### READ AND CALIBRATE NECESSARY DATASETS ###
    ## Read proxy Data
    reg_data <- get_proxy(start_72, end_72, pollutant)
    
    ## Read AQY data
    aqy_data_path <- list.files('C:/Users/18313/Desktop/airMonitoring/downloader/results/concatenated_data', '*freq_60', full.names = T)
    aqy_data_full <- read.csv(aqy_data_path[length(aqy_data_path)], stringsAsFactors = F) %>%
      rename('timestamp'='Time')
    
      # Calibrate AQY data
      cal_files_O3_all <- list.files('results', 'cal_params_O3*', full.names = T)
      cal_file_O3 <- read.csv(cal_files_O3_all[length(cal_files_O3_all)], stringsAsFactors = F) %>%
        filter(!is.na(lat))
      
      cal_files_NO2_all <- list.files('results', 'cal_params_NO2*', full.names = T)
      cal_file_NO2 <- read.csv(cal_files_NO2_all[length(cal_files_NO2_all)], stringsAsFactors = F) %>%
        filter(!is.na(lat))
      
      aqy_data_O3 <- inner_join(aqy_data_full, cal_file_O3, by = 'ID') %>% # Join data with calibration parameters, filter out non-applicable parameters
        filter(timestamp >= start_date & timestamp <= end_date & timestamp >= start_72 & timestamp <= end_72) %>%
        mutate(O3_cal = O3.offset + O3.gain*O3)
      
      aqy_data_NO2 <- dplyr::select(aqy_data_O3, ID, timestamp, O3_cal, Ox, NO2) %>%
        inner_join(cal_file_NO2, by = 'ID') %>%
        filter(timestamp >= start_date & timestamp <= end_date & timestamp >= start_72 & timestamp <= end_72) %>%
        mutate(NO2_cal = NO2.b0 + NO2.b1*Ox + NO2.b2*O3_cal)
            
    ### DETECT DRIFT
      aqy_data_72 <- switch(pollutant, 'OZONE' = aqy_data_O3, 'NO2' = aqy_data_NO2, stop('Error: accepts inputs of OZONE or NO2.'))
      aqy_data_72$aqy_raw <- switch(pollutant, 'OZONE' = aqy_data_72$O3, 'NO2' = aqy_data_72$NO2)
      aqy_data_72$aqy_cal <- switch(pollutant, 'OZONE' = aqy_data_72$O3_cal, 'NO2' = aqy_data_72$NO2_cal)
    
      ## Specify monitors for drift detection and initialize results
        aqy_list_id <- unique(pull(aqy_data_72, ID)) 
          
        todays_flags <- as.data.frame(aqy_list_id) %>%
          mutate(ks = 0, gain = 0, offset = 0) %>%
          rename('ID'='aqy_list_id')
        
      ## Generate flags by monitor
        for(i in 1:length(aqy_list_id)){
                  
          # Filter to the AQY we would like to test for drift
          aqy_ID <- aqy_list_id[i]
          aqy_for_cal <- filter(aqy_data_72, ID == aqy_ID)
                        
          # Filter to appropriate proxy
          proxy_stn <- unique(aqy_for_cal$proxy_site)
          print(proxy_stn)
          proxy_for_cal <- subset(reg_data, reg_data$proxy_site %in% proxy_stn)
                            
          # Test for monitor drift if sufficient data in AQY and proxy data sets
          if(nrow(aqy_for_cal) <= .75*nrow(proxy_for_cal) | nrow(aqy_for_cal)*.75 >= nrow(proxy_for_cal)){
            print(paste('Did not perform', pollutant, 'drift detection for', aqy_ID, 'due to insufficient data. NROW AQY = ', nrow(aqy_for_cal), 'NROW Proxy =', nrow(proxy_for_cal)))
            }
                  
            else{ 
              # Kolmogorov-Smirnov test
              ks_results <- ks.test(proxy_for_cal$proxy_rand, aqy_for_cal$aqy_cal, exact = F)
              ks_p <- ks_results$p.value # Get p-value from KS test
                                
              todays_flags$ks[i] <-ifelse(ks_p <= 0.05, 1, 0)
                                
              # Mean-variance moment matching for gain
              manual_gain <- sqrt(var(proxy_for_cal$proxy_rand, na.rm = T) / var(aqy_for_cal$aqy_cal, na.rm = T))
              todays_flags$gain[i] <- ifelse(manual_gain > 1.3 | manual_gain < .7, 1, 0)
                                
              # Mean-variance moment matching for offset
              manual_offset <- mean(proxy_for_cal$proxy_rand, na.rm = T) - mean(aqy_for_cal$aqy_cal, na.rm = T)*manual_gain
              todays_flags$offset[i] <- ifelse(manual_offset > 5 | manual_offset < -5, 1, 0) # Assign flag if manual offset outside bounds
                                
              # Print results
              print(paste(aqy_ID, 'from', start_72, 'to', end_72, 'for', pollutant, '|| KS P-VALUE:', ks_p, '| MANUAL GAIN:', manual_gain, '| MANUAL OFFSET:', manual_offset))
              }
            }
              
            # Combine today's flags with existing flags
            running_flags <- read.csv(paste0('results/running_flags_', pollutant, '.csv'), stringsAsFactors = F) %>% 
              full_join(., todays_flags, by = 'ID') %>% 
              mutate(ks_run = ifelse(ks == 0, 0, ks_run + ks), gain_run = ifelse(gain == 0, 0, gain_run + gain), offset_run = ifelse(offset == 0, 0, offset_run + offset)) %>%
              dplyr::select(-c(ks, gain, offset))
              
              write.csv(running_flags, paste0('results/running_flags_', pollutant, '.csv'), row.names = F)
        
    ### RECALIBRATE ###
      ## Set important dates
      start_recal <- end_72-8 
      end_recal <- end_72 
      flag_detected <- end_72-5
      
      ## Identify monitors needing re-calibration, using highest of flagged columns
      running_flags$max_flag <- apply(MARGIN = 1, X = running_flags[grep('*_run', colnames(running_flags))], FUN = max)
      needs_calibration <- pull(filter(running_flags, max_flag >= 5), ID)
      
      ## Re-calibrate if running flags >= 5 for any monitors
      if(length(needs_calibration) == 0){print(paste('No recalibration is needed for period from', start_72, 'to', end_72, 'for', pollutant))}
    
        else{# Re-calibrate using last eight days of data
            aqy_recal <- switch(pollutant, 
                                'OZONE' = filter(aqy_data_O3, ID %in% needs_calibration & timestamp >= start_recal & timestamp <= end_recal), 
                                'NO2' = filter(aqy_data_NO2, ID %in% needs_calibration & timestamp >= start_recal & timestamp <= end_recal),
                                stop(print('Error: accepts input of OZONE or NO2.')))
                         
            proxy_recal <- get_proxy(start_recal, end_recal, pollutant)
            
            # Blank data frame for new parameters
            proxy_sites_recal <- unique(as.data.frame(cbind(ID = aqy_recal$ID, proxy_site = aqy_recal$proxy_site)))
            
            new_params <- as.data.frame(needs_calibration) %>%
              rename('ID'='needs_calibration') %>%
              left_join(proxy_sites_recal, by = 'ID') %>%
              mutate(O3.gain = 1, O3.offset = 0, NO2.b0 = 0, NO2.b1 = 1, NO2.b2 = 1, NO2.b0_kl = 0, NO2.b1_kl = 1, NO2.b2_kl =1,  
                     start_date = as.character(flag_detected), end_date = '9999-12-31')

            for(i in 1:nrow(new_params)){
              
              # Isolate data for desired AQY & proxy site
              aqy_recal_i <- filter(aqy_recal, ID == new_params$ID[i])
              proxy_recal_i <- filter(proxy_recal, proxy_site == new_params$proxy_site[i])
              recal_i <- na.omit(inner_join(aqy_recal_i, proxy_recal_i, 'timestamp'))

              if(pollutant == 'OZONE'){
                # Calculate new gain and offset for O3
                gain_new_O3 <- sqrt(var(proxy_recal_i$proxy_rand, na.rm = T)/var(aqy_recal_i$O3, na.rm = T))
                offset_new_O3 <- mean(proxy_recal_i$proxy_rand, na.rm = T) - gain_new_O3*mean(aqy_recal_i$O3, na.rm = T)
                
                new_params$O3.gain[i] <- gain_new_O3
                new_params$O3.offset[i] <- offset_new_O3}
              
              else{
                # Calculate new gain and offset for NO2 using mean-variance moment matching
                new_params$NO2.b0[i] <- mean(proxy_recal_i$proxy_rand, na.rm = T) - mean(aqy_recal_i$Ox - aqy_recal_i$O3_cal, na.rm = T)
                new_params$NO2.b1[i] <- sqrt(var(proxy_recal_i$proxy_rand, na.rm = T)/var(aqy_recal_i$Ox - aqy_recal_i$O3_cal, na.rm = T))
                new_params$NO2.b2[i] <- new_params$NO2.b1[i]
                
                # Calculate new gain and offset for NO2 using objective function minimization
                b0 <- Variable(1)
                b1 <- Variable(1)
                b2 <- Variable(1)
  
                cno2 <- b0 + b1*recal_i$Ox - b2*recal_i$O3_cal
                pno2 <- recal_i$proxy_raw
    
                kl <- kl_div(cno2, pno2)
                kl_obj <- sum(kl)
    
                kl_min <- Problem(objective = Minimize(kl_obj))
                kl_out <- solve(kl_min)
    
                new_params$NO2.b0_kl[i] <- kl_out$getValue(b0)
                new_params$NO2.b1_kl[i] <- kl_out$getValue(b1)
                new_params$NO2.b2_kl[i] <- kl_out$getValue(b2)
                }
          }
        
        ## Work new parameters into running list
          # Edit old parameters to combine with new
          params_existing <- switch(pollutant, 'OZONE' = cal_file_O3, 'NO2' = cal_file_NO2)
            
          params_existing$end_date <- ifelse(params_existing$ID %in% new_params$ID & params_existing$end_date == '9999-12-31', # Set new start and end dates for re-calibrated AQYs
                                             as.character(as.Date(new_params$start_date)-1), 
                                             params_existing$end_date)
          
          # Edit new parameters to combine with old
          desired_cols <- colnames(params_existing)
          
          params_current <- dplyr::select(params_existing, c('ID', 'description', 'address', 'city', 'lat', 'lon')) %>%
            right_join(., new_params, by = 'ID') %>%
            dplyr::select(desired_cols)
          
          # Write combined parameters as output
          running_params <- rbind(params_current, params_existing) %>%
            unique() %>%
            arrange(ID, start_date)
        
          write.csv(running_params, paste0('results/cal_params_', pollutant, '_', end_recal, '.csv'), row.names = F)
          
          # Reset running flags for re-calibrated data
          mutate(running_flags, ks_run = ifelse(ID %in% needs_calibration, 0, ks_run),
                                gain_run = ifelse(ID %in% needs_calibration, 0, gain_run), 
                                offset_run = ifelse(ID %in% needs_calibration, 0, offset_run)) %>%
            dplyr::select(-max_flag) %>%
            write.csv(paste0('results/running_flags_', pollutant, '.csv'), row.names = F)
                    
        }
}
```

### RUN FUNCTION FOR ALL DAYS BACK TO LAST RE-CALIBRATION
```{r}
O3_recals <- list.files('results', 'cal_params_OZONE*')
O3_last_recal <- str_extract(O3_recals[length(O3_recals)], '\\d\\d\\d\\d-\\d\\d-\\d\\d')
O3_to_run <- seq(as_date(O3_last_recal), Sys.Date()-3, 1)

NO2_recals <- list.files('results', 'cal_params_NO2*')
NO2_last_recal <- str_extract(NO2_recals[length(NO2_recals)], '\\d\\d\\d\\d-\\d\\d-\\d\\d')
NO2_to_run <- seq(as_date(NO2_last_recal), Sys.Date()-3, 1)

# Run ozone first
for(i in O3_to_run){calibrate_monitors(as_date(i), 'OZONE')}

# Run NO2 next
for(i in NO2_to_run){calibrate_monitors(as_date(i), 'NO2')}
```


































