---
title: "Richmond Network Complete Calibration Framework"
author: "Audrey Smith"
date: "5/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(httr)
library(jsonlite)
library(dplyr)
library(tidyr)
library(lubridate)
library(stringr)
library(CVXR)
```

# Basic Setup
```{r}
td_hour <- 60*60
td_day <- td_hour*24
td_30day <- td_day*30
td_72hr <- 60*60*72
td_5day <- 60*60*24*5
```

```{r}
format_timestamp <- function(timestamp){
  
  formatted_timestamp <- as.character(format(as.POSIXct(timestamp), '%Y-%m-%d %H:%M:%S'))
  
  return(formatted_timestamp)
}
```

```{r}
reset_flags <- function(pollutant){
  
  file_with_flags <- paste0('results/running_flags/running_flags_', pollutant, '.csv')
  
  read.csv(file_with_flags) %>%
    mutate_at(.vars = c('ks', 'gain', 'offset'), .funs = ~.x*0) %>%
    write.csv(file_with_flags, row.names = F)
  
  print(paste('Reset', pollutant, 'flags - restarting drift detection at zero'))
}
```


# Get Proxy Data
```{r}
get_existing_proxy_data <- function(pollutant){
  
  existing_proxy_data_path <- paste0('results/proxy/', pollutant, '_proxy_data.csv')
  
  try(existing_proxy_data <- read.csv(existing_proxy_data_path, stringsAsFactors = F))
    
  if(!exists('existing_proxy_data', where = environment())){existing_proxy_data <- data.frame()
      print(paste('No existing proxy data for', pollutant, '- creating new'))}
      
    else(print(paste('Reading existing proxy data from', existing_proxy_data_path)))

  return(existing_proxy_data)
  
}
```

```{r}
find_starttime_proxy_request <- function(existing_proxy_data){
  
  ramn_start_time <- format_timestamp('2019-12-01 00:00:00')
  yesterday <- format_timestamp(with_tz(Sys.time() - td_day, 'UTC'))
  right_now <- format_timestamp(with_tz(round(Sys.time(), 'hour'), 'UTC'))
  
  if(nrow(existing_proxy_data) > 0){timestamps_existing <- unique(existing_proxy_data$timestamp_utc)}
    else{print('No existing data - requesting all starting Dec 2019')
        return(ramn_start_time)}
  
  last_timestamp_recorded <- max(timestamps_existing)
  
  next_timestamp_needed <- format_timestamp(ymd_hms(last_timestamp_recorded) + td_hour)
  
  print(paste('Requesting data starting with', next_timestamp_needed))
  
  return(next_timestamp_needed)
  
}
```

```{r}
build_proxy_request_url <- function(pollutant, start_request){
  
  month_after_start <- suppressWarnings(as.character(ymd_hms(start_request) + td_30day))
  if(is.na(month_after_start)){month_after_start <- paste(as.character(as.Date(start_request) + td_30day), '00:00:00')}

  yesterday <- as.character(Sys.time() - 60*60*24)
  
  end_request <- min(c(month_after_start, yesterday))
  
  first_date_needed <- as.Date(start_request)
  first_hour_needed <- hour(start_request)
          
  last_date_needed <- as.Date(end_request)
  last_hour_needed <- hour(end_request)
          
  request_url = paste0('http://www.airnowapi.org/aq/data/?',
                        'startDate=', first_date_needed, 'T', first_hour_needed, '&endDate=', last_date_needed, 'T', last_hour_needed, 
                        '&parameters=', pollutant,
                        '&BBOX=-122.617880,37.639710,-121.706015,38.177130',
                        '&dataType=C&format=application/json&verbose=1&nowcastonly=0&includerawconcentrations=1',
                        '&API_KEY=C05358E3-5508-4216-A03E-E229E0368B7E')
          
   print(paste('Requesting proxy data from Air District for', pollutant, 'from', first_date_needed, first_hour_needed, 'to', last_date_needed, last_hour_needed))
  
   return(request_url)
}
```

```{r}
request_new_proxy_data <- function(request_url){
  
  request_call <- GET(url = request_url)
  stop_for_status(request_call, paste('Failed request to URL', request_url))
  
  request_json <- content(request_call, as = 'text', type = NULL, encoding = 'UTF-8')
  request_df <- fromJSON(request_json, simplifyDataFrame = TRUE)
  
  new_proxy_data <- mutate(request_df, 
                           timestamp_utc = paste0(substr(UTC, 1, 10), ' ', substr(UTC, 12, 16), ':00'), 
                           RawConcentration = ifelse(RawConcentration == -999, Value, RawConcentration))
  
  return(new_proxy_data)
  
}
```

```{r}
clean_new_proxy_data <- function(new_proxy_data){

   random_noise <- runif(nrow(new_proxy_data), .000001, .0001)
   new_proxy_data$proxy_rand <- new_proxy_data$RawConcentration + random_noise
    
   new_proxy_clean <- dplyr::select(new_proxy_data, -c(UTC, Value, AgencyName, FullAQSCode, IntlAQSCode)) %>% 
     rename('lat'='Latitude', 'long'='Longitude', 'modality'='Parameter', 'units'='Unit', 'proxy_raw'='RawConcentration', 'proxy_site' = 'SiteName')

   return(new_proxy_clean)
}
```

```{r}
request_all_dates_needed <- function(pollutant, request_start_time, existing_proxy_data){
  
  while(request_start_time < as.character(Sys.time() - td_day)){  
    
      request_url <- build_proxy_request_url(pollutant, request_start_time)
      
      new_proxy_data_raw <- request_new_proxy_data(request_url)
      new_proxy_data <- clean_new_proxy_data(new_proxy_data_raw)
      
      existing_proxy_data <- rbind(existing_proxy_data, new_proxy_data)
      
      month_after_last_start <- format_timestamp(ymd_hms(request_start_time) + td_30day + td_hour)
      
      request_start_time <- min(format_timestamp(Sys.time() - td_day), month_after_last_start)
  }
  
  return(existing_proxy_data)
}
```

```{r}
process_and_write_proxy_data <- function(pollutant, proxy_data){
  
  proxy_data_no_duplicates <- group_by(proxy_data, proxy_site, timestamp_utc) %>%
    slice_sample()
  
  print(paste('Proxy data for', pollutant, 'now current - writing to results/proxy'))
  
  write.csv(proxy_data_no_duplicates, paste0('results/proxy/', pollutant, '_proxy_data.csv'), row.names = F)
  
  return(proxy_data_no_duplicates)
}
```

_Final Proxy Function_
```{r}
get_current_proxy_data <- function(pollutant){
  
  existing_proxy_data <- get_existing_proxy_data(pollutant)

  request_start_time <- find_starttime_proxy_request(existing_proxy_data)

  updated_proxy_data_raw <- request_all_dates_needed(pollutant, request_start_time, existing_proxy_data)
  
  updated_proxy_data <- process_and_write_proxy_data(pollutant, existing_proxy_data)
  
  return(updated_proxy_data)
  
}
```

####################################################

# Get AQY Data
```{r}
read_aqy_data <- function(){
  
  files_at_path <- list.files(path = 'raw_data', pattern = '*.csv', full.names = T)
  
  print(files_at_path)
  
  if(length(files_at_path) < 1){stop('No existing AQY data found in raw data directory - please add hourly data')}
    else if(length(files_at_path) > 1){stop('Too many files in raw data directory - please only include hourly data')}
      else{aqy_data <- read.csv(files_at_path[1], stringsAsFactors = F)}
  
  aqy_data_utc <- aqy_data %>% 
    rename('timestamp_original'='Time') %>%
    mutate(timestamp_utc = ymd_hms(timestamp_original, tz = 'America/Los_Angeles'),
           timestamp_utc = with_tz(timestamp_utc, tz = 'UTC'),
           timestamp_utc = as.character(timestamp_utc), .after = 'ID') %>%
    filter(!is.na(timestamp_utc))
  
  return(aqy_data_utc)
}
```

```{r}
get_current_params <- function(pollutant){
  
  all_params_paths <- list.files('results', paste0(pollutant, '*'), full.names = T)
  current_params_path <- all_params_paths[length(all_params_paths)]
  current_params <- read.csv(current_params_path)
  
  print(paste('Calibrating with most recent', pollutant, 'params at', current_params_path))
  
  return(current_params)
  
}
```

```{r}
calibrate_aqy_data <- function(aqy_data, pollutant){
  
  params_O3 <- get_current_params('OZONE')
  params_NO2 <- get_current_params('NO2')
  
  calibrated_data_O3 <- inner_join(params_O3, aqy_data, by = 'ID') %>%
    filter(timestamp_original >= deployment_datetime & timestamp_utc >= start_date & timestamp_utc <= end_date) %>%
    mutate(O3 = O3.offset + (O3.gain*O3_raw)) %>%
    dplyr::select(ID, timestamp_utc, timestamp_original, TEMP, RH, DP, PM25_raw, O3, O3_raw, Ox_raw, NO2_raw)
  
  calibrated_data <- inner_join(params_NO2, calibrated_data_O3, by = 'ID') %>%
    mutate(NO2 = NO2.offset + (NO2.gain*NO2_raw), .before = 'NO2_raw')
  
  pollutant_name <- switch(pollutant, 'OZONE'='O3', 'NO2'='NO2')
  
  new_col_names <- str_replace_all(string = colnames(calibrated_data), pattern = pollutant_name, replacement = 'pollutant')
  colnames(calibrated_data) <- new_col_names
  
  return(calibrated_data)
  
}
```

```{r}
filter_and_join_data <- function(aqy_data, proxy_data, start_time, td_add_to_start){
  
  end_time <- format_timestamp(ymd_hms(start_time) + td_add_to_start)
  
  print(paste('Detecting drift for data between', start_time, 'and', end_time))
  
  joined_data <- inner_join(aqy_data, proxy_data, by = c('timestamp_utc', 'proxy_site')) %>%
    filter(ID %in% c('AQY BB-633', 'AQY BB-642') & timestamp_utc < '2021') %>%
    filter(timestamp_utc < end_time) %>%
    na.omit()
  
  return(joined_data)
  
}
```

#######################################


# Detect Drift
```{r}
detect_drift <- function(joined_data_72hr){
          
  drift_tests <- group_by(joined_data_72hr, ID) %>%
    summarize(manual_gain = sqrt(var(proxy_rand)/var(pollutant)),
              manual_offset = mean(proxy_rand) - manual_gain*mean(pollutant),
              ks_p = suppressWarnings(ks.test(proxy_rand, pollutant, exact = F))$p.value, 
              .groups = 'drop_last')
  
  drift_flags <- mutate(drift_tests,
                        gain_new = ifelse(manual_gain >= 1.3 | manual_gain <= .7, 1, 0),
                        offset_new = ifelse(manual_offset >= 5 | manual_offset <= -5, 1, 0),
                        ks_new = ifelse(ks_p <= .05, 1, 0))
                        # add test for extreme values relative to rest of data distribution
  
  # To delete print statements after build out done - simply a test
  results_633 <- filter(drift_tests, ID == 'AQY BB-633')
  results_642 <- filter(drift_tests, ID == 'AQY BB-642')
  print(paste('Results for 633: ks pval =', results_633$ks_p, '| gain =', results_633$manual_gain, '| offset =', results_633$manual_offset))
  print(paste('Results for 642: ks pval =', results_642$ks_p, '| gain =', results_642$manual_gain, '| offset =', results_642$manual_offset))
          
  new_flags <- dplyr::select(drift_flags, -c('manual_gain', 'manual_offset', 'ks_p')) %>%
    mutate_at(.vars = c('ks_new', 'gain_new', 'offset_new'), .funs = as.integer)
  
  aqys_with_flags <- filter(new_flags, gain_new > 0 | offset_new > 0 | ks_new > 0) %>%
    pull(ID)
  
  if(length(aqys_with_flags) > 0){print(paste(aqys_with_flags, 'flagged for this period'))}
    else{print('No AQYs flagged for this period')}
      
  return(new_flags)
}
```

```{r}
sum_old_and_new_flags <- function(pollutant, new_flags){
  
  flag_path <- paste0('results/running_flags/running_flags_', pollutant, '.csv')
  existing_flags <- read.csv(flag_path, stringsAsFactors = F)

  joined_flags <- full_join(existing_flags, new_flags, by = 'ID')

  cols_to_rmv_na <- colnames(joined_flags)[2:ncol(joined_flags)]
  
  sum_if_flagged <- function(old, new){ifelse(new == 0, 0, old + new)}
  
  summed_flags <- mutate_at(.tbl = joined_flags, .vars = cols_to_rmv_na, .funs = ~replace_na(.x, 0)) %>%
    transmute(ID = ID, 
              ks = sum_if_flagged(ks, ks_new),
              gain = sum_if_flagged(gain, gain_new),
              offset = sum_if_flagged(offset, offset_new))
  
  write.csv(summed_flags, flag_path, row.names = F)
  
  return(summed_flags)
}
```

```{r}
get_aqys_needing_recal <- function(summed_flags){
  
  largest_running_flag <- apply(X = summed_flags[2:ncol(summed_flags)], MARGIN = 1, FUN = max)
  
  summed_flags$max_flag <- largest_running_flag
  
  largest_flag_network <- max(largest_running_flag)
  
  print(paste('Largest flag:', largest_flag_network))
  
  aqys_needing_recal <- filter(summed_flags, max_flag >= 24*5) %>%
    pull(ID)
  
  return(aqys_needing_recal)
  
}
```

#######################################

# Recalibrate Data
```{r}
generate_new_gain_and_offset <- function(joined_data_30day, aqys_needing_recal, pollutant){
  
  new_gain_and_offset <- filter(joined_data_30day, ID %in% aqys_needing_recal) %>%
    group_by(ID) %>%
    summarize(new.gain = sqrt(var(proxy_rand)/var(pollutant_raw)),
              new.offset = mean(proxy_rand) - new.gain*mean(pollutant_raw),
              .groups = 'drop_last')
  
  new_col_names <- switch(pollutant, 
                          'OZONE'=c('ID', 'O3.gain', 'O3.offset'), 
                          'NO2'=c('ID', 'NO2.gain', 'NO2.offset'))
  colnames(new_gain_and_offset) <- new_col_names
  
  print(as.matrix(new_gain_and_offset))

  return(new_gain_and_offset)
  
  }
```

```{r}
combine_new_and_old_params <- function(new_params, pollutant, start_time_rolling_72hr){
  
  old_params_path_all <- list.files('results', paste0(pollutant, '*'), full.names = T, recursive = F)
  old_params_path_newest <- old_params_path_all[length(old_params_path_all)]
  old_params <- read.csv(old_params_path_newest)
  
  aqy_metadata <- select(old_params, c('ID', 'deployment_date', 'deployment_datetime', 'Longitude', 'Latitude', 'proxy_site')) %>%
    unique()
  
  time_first_flagged <- start_time_rolling_72hr - td_5day
  
  new_params_metadata <- inner_join(aqy_metadata, new_params, by = 'ID') %>%
    mutate(start_date = format_timestamp(time_first_flagged), end_date = '2199-12-31 23:59:59', .after = proxy_site)

  combined_params <- rbind(old_params, new_params_metadata) %>%
    arrange(ID, start_date) %>%
    group_by(ID) %>%
    mutate(end_date = ifelse(end_date == '2199-12-31 23:59:59' & start_date != max(start_date), 
                             format_timestamp(ymd_hms(max(start_date)) - td_hour), 
                             end_date))

  timestamp_for_path <- str_remove_all(format_timestamp(time_first_flagged), '\\:')
  new_param_path <- paste0('results/', pollutant, '_calvals_', timestamp_for_path, '.csv')
  write.csv(combined_params, new_param_path, row.names = F)
  
  return(time_first_flagged)
}
```

```{r}
reset_flags_if_recalibrated <- function(pollutant){
  
  all_flags <- read.csv(paste0('results/running_flags/running_flags_', pollutant, '.csv'))
  
  flags_to_reset <- filter(all_flags, ks >= 120 | gain >= 120 | offset >= 120)
  aqys_to_reset <- unique(flags_to_reset$ID)
  flags_cleared <- mutate_at(flags_to_reset, .vars = c('ks', 'gain', 'offset'), .funs = ~.x*0)
  
  flags_to_leave <- filter(all_flags, ID %in% aqys_to_reset == F)
  
  updated_flags <- rbind(flags_to_leave, flags_cleared) %>%
    arrange(ID)
  
  write.csv(updated_flags, paste0('results/running_flags/running_flags_', pollutant, '.csv'), row.names = F)
  
  print('Reset flags - restarting drift detection at zero')
}
```

### Function to Combine Drift Detection and Calibration
```{r}
check_drift_calibrate_data <- function(aqy_data, proxy_data, pollutant, start_time_rolling_72hr){
  
  aqy_data <- calibrate_aqy_data(aqy_data, pollutant)
  
  start_time_recalibration <- as.character(ymd_hms(start_time_rolling_72hr) - td_30day)
  joined_data_72hr <- filter_and_join_data(aqy_data, O3_proxy, start_time_rolling_72hr, td_72hr)
  
  print(paste('Performing drift detection for', start_time_rolling_72hr, 'with', nrow(joined_data_72hr), 'rows of data'))
  
  flags_72hr <- detect_drift(joined_data_72hr)
  summed_flags <- sum_old_and_new_flags(pollutant, flags_72hr)
  aqys_needing_recal <- get_aqys_needing_recal(summed_flags)
  
  if(length(aqys_needing_recal) < 1){return(print('No monitors require recalibration for this 72-hr period'))}
    else{
      print('Recalibrating monitors')
  
      joined_data_30day <- filter_and_join_data(aqy_data, O3_proxy, start_time_recalibration, td_30day)
      new_params <- generate_new_gain_and_offset(joined_data_30day, aqys_needing_recal, pollutant)
      combined_params <- combine_new_and_old_params(new_params, pollutant, start_time_rolling_72hr)
      
      reset_flags_if_recalibrated(pollutant)
      
      return(combined_params)
    }
}
```

#######################################


# Run Everything
#######################################

```{r}
O3_proxy <- get_current_proxy_data('OZONE')
NO2_proxy <- get_current_proxy_data('NO2')
PM25_proxy <- get_current_proxy_data('PM25')

aqy_data <- suppressWarnings(read_aqy_data())
```


reset_flags('NO2')

timestamps <- filter(aqy_data, ID %in% c('AQY BB-633', 'AQY BB-642')) %>%
  pull(timestamp_utc) %>%
  unique()

for(timestamp in timestamps){
  
  print(timestamp)
  
  check_drift_calibrate_data(aqy_data, O3_proxy, 'NO2', format_timestamp(timestamp))













recalibrate_NO2 <- function(aqy_id, thirty_day_aqy, thirty_day_proxy){
  
  # Generate data
  thirty_day_aqy_filter <- filter(thirty_day_aqy, ID == aqy_id)
  thirty_day_proxy_filter <- filter(thirty_day_proxy, proxy_site == unique(thirty_day_aqy_filter$proxy_site))
  thirty_day_data <- na.omit(inner_join(thirty_day_aqy_filter, thirty_day_proxy_filter, 'timestamp'))
  
  # Mean-variance moment matching
  NO2.b0_byhand <- mean(thirty_day_data$proxy_rand) - mean(thirty_day_data$Ox - thirty_day_data$O3_cal)
  NO2.b1_byhand <- sqrt(var(thirty_day_data$proxy_rand)/var(thirty_day_data$Ox - thirty_day_data$O3_cal))
  NO2.b2_byhand <- NO2.b1_byhand
  
  #Objective function minimization
  ### TO DO - FILL IN BLANK IF TRY FAILS
  ### TO DO - LOOK INTO SOLVER ABILITY TO INITIALIZE WITH SPECIFIED VALUES
  try({
    b0 <- Variable(1)
    b1 <- Variable(1)
    b2 <- Variable(1)

    cno2 <- recal_filter$Ox - b1*recal_filter$O3_cal
    pno2 <- recal_filter$proxy

    kl <- kl_div(pno2, cno2)
    kl_obj <- sum(kl)
    
    objective <- Minimize(kl_obj)
    constraints <- list(cno2 >= 0)

    kl_min <- Problem(objective, constraints)
    kl_out <- solve(kl_min)
    
    NO2.b0_kl <- kl_out$getValue(b0)
    NO2.b1_kl <- kl_out$getValue(b1)
    NO2.b2_kl <- kl_out$getValue(b2)
  })
  
  warning('Objective function solver failure: defaulting to hand calculated values' = !exists('NO2.b0_kl'))
  
  NO2.b0 <- if(exists('NO2.b0_kl')){NO2.b0_kl}
                else{NO2.b0_byhand}
  
  NO2.b1 <- if(exists('NO2.b1_kl')){NO2.b1_kl}
                else{NO2.b1_byhand}
  
  NO2.b2 <- if(exists('NO2.b2_kl')){NO2.b2_kl}
                else{NO2.b2_byhand}
  
  #Combined results
  NO2_params <- as.data.frame(cbind(aqy_id, NO2.b0, NO2.b1, NO2.b2))
  
  print(paste0(aqy_id, ': ', 'new b0', NO2.b0, ', new b1 ', NO2.b1, ', new b2', NO2.b2))
  
  return(NO2_params)
}
         































